---
title: "Practical Machine Learning Course Project"
author: "Cai Jiashen"
date: "8/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Practical Machine Learning Course Project
## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

## Task
Create prediction models incorporating various measurements/data from personal activity devices to predict performance in doing Unilateral Dumbbell Biceps Curls.

## Loading Library
Loading required libraries.
```{r message=FALSE, warning=FALSE, results=FALSE}
library(caret)
library(dplyr)
library(parallel)
library(doParallel)
library(rattle)
library(ggplot2)
```

## Loading Data Sets
Loading datasets and setting up factors.
```{r}
pml_train <- read.csv("pml-training.csv", na.strings=c("","NA"))
pml_test <- read.csv("pml-testing.csv", na.strings=c("","NA"))

pml_train$user_name <- factor(pml_train$user_name)
pml_test$user_name <- factor(pml_test$user_name)
pml_train$classe <- factor(pml_train$classe)

str(pml_train)

```
## Data Cleaning and Slicing
Cleaning up data - Removing empty variables ("" or "NA), variables with near zero variance.
```{r}
pml_train_clean <- select(pml_train, -c(X, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp))
NZV <- nearZeroVar(pml_train_clean)
pml_train_clean <- pml_train_clean[,-NZV]
pml_train_clean <- pml_train_clean[,colSums(is.na(pml_train_clean)) == 0]

str(pml_train_clean)

inTrain <- createDataPartition(pml_train_clean$classe, p = 0.7, list = FALSE)
trainSet <- pml_train_clean[inTrain,]
testSet <- pml_train_clean[-inTrain,]
```

## Prediction Model
### Setup Parallel Computing
```{r}
cluster <- makeCluster(detectCores() - 1, setup_timeout = 0.5) # convention to leave 1 core for OS
registerDoParallel(cluster)
fitControl <- trainControl(method = "cv", number = 3, allowParallel = TRUE, verboseIter = TRUE)
```

### Decision Tree Model
```{r}
set.seed(2020)
modFitDT <- train(classe ~ ., data=trainSet, method="rpart", trControl=fitControl)
modFitDT
fancyRpartPlot(modFitDT$finalModel)

predictDT <- predict(modFitDT, newdata = testSet)
confMatrixDT <- confusionMatrix(predictDT, testSet$classe)
```

### Random Forest Model
```{r}
set.seed(2020)
modFitRF <- train(classe ~ ., data=trainSet, method="rf", trControl=fitControl)
modFitRF

predictRF <- predict(modFitRF, newdata = testSet)
confMatrixRF <- confusionMatrix(predictRF, testSet$classe)
```

### Gradient Boosted Model
```{r}
set.seed(2020)
modFitGBM <- train(classe ~ ., data=trainSet, method="gbm", trControl=fitControl, verbose = FALSE)
modFitGBM

predictGBM <- predict(modFitGBM, newdata = testSet)
confMatrixGBM <- confusionMatrix(predictGBM, testSet$classe)
```
### Stop Parallel Computing
```{r}
stopCluster(cluster)
registerDoSEQ()
```

## Summarise Prediction Models
Summary of Prediction Models in terms of Accuracy and Kappa statistic.
```{r}
modelSummary <- data.frame(Model =c("Decision Tree", "Random Forest", "Gradient Boosting"), Accuracy = c(confMatrixDT$overall[1], confMatrixRF$overall[1], confMatrixGBM$overall[1]), Kappa = c(confMatrixDT$overall[2], confMatrixRF$overall[2], confMatrixGBM$overall[2]))
modelSummary
ggplot(data = modelSummary, aes(x = Kappa, y = Accuracy, colour = Model)) +
  geom_point() +
  xlim(0, 1) + ylim (0, 1) +
  ggtitle("Accuracy and Kappa of Prediction Models")
```

## Apply Best Model to Test Data
Using the Random Forest model as it offers the best Accuracy and Kappa statistic.
```{r}
predictTest <- predict(modFitRF, newdata=pml_test)
predictTest

```
